---
title: "MiMIR report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    theme: paper
always_allow_html: true
params:
  metabo_measures: metabo_measures
  phenotypes: phenotypes
  bin_phenotypes: bin_phenotypes
  bin_pheno_available: bin_pheno_available
  mort_score: mort_score
  MetaboAge: MetaboAge
  surrogates: surrogates
  predictors: predictors
  calibrations: calibrations
  Nbins: Nbins
  Nmax_miss_metaboAge: Nmax_miss_metaboAge
  Nmax_zero_metaboAge: Nmax_zero_metaboAge
  Nmax_miss_surrogates: Nmax_miss_surrogates
  Nmax_zero_surrogates:  Nmax_zero_surrogates
---

# Reading packages and variables
Required packages are loaded and the input variables are set. The used input settings are shown.
```{r setup}
if (!require("knitr")) install.packages("knitr")
if (!require("tidyr")) install.packages("tidyr")
if (!require("plotly")) install.packages("plotly")
if (!require("plyr")) install.packages("plyr")
```

```{r variables,echo=FALSE}
# IMPORT VARIABLES
metabo_measures<-params$metabo_measures
phenotypes<-params$phenotypes
bin_phenotypes<-params$bin_phenotypes
bin_pheno_available<-params$bin_pheno_available
mort_score<-params$mort_score
MetaboAge<-params$MetaboAge
surrogates<-params$surrogates
predictors<-params$predictors
calibrations<-params$calibrations
Nbins<-params$Nbins
Nmax_miss_metaboAge<-params$Nmax_miss_metaboAge
Nmax_zero_metaboAge<-params$Nmax_miss_metaboAge
Nmax_miss_surrogates<-params$Nmax_miss_metaboAge
Nmax_zero_surrogates<-params$Nmax_miss_metaboAge
```

# Dataset Characteristics
### Dimensions of the files
```{r showVariables, echo=FALSE}
cat(paste0("Metabolites file dimensions: \nRows= ", dim(metabo_measures)[1], ", Columns= ", dim(metabo_measures)[2]),"\nAll the necessary metabolites were found available!",
    "\n\nPhenotypes file dimensions: \nRows= ", dim(phenotypes)[1], ", Columns= ", dim(phenotypes)[2])
```

## Metabolites
### Missingness of the metabolites
```{r missing metabolites, echo=FALSE}
plot_na_heatmap(t(metabo_measures[,MiMIR::metabolites_subsets$MET57]))
```

### Correlation of the metabolites 
```{r corr metabolites, echo=FALSE}
  res<-cor_assoc(metabo_measures,metabo_measures, MiMIR::metabolites_subsets$MET57, MiMIR::metabolites_subsets$MET57)
  plot_corply(res, main="Metabolites' Correlations", reorder.x=TRUE, abs=F, 
                      resort_on_p= TRUE,reorder_dend=F)
```

### Histogram metabolites
```{r histograms metabolites, echo=FALSE, out.width="150%", out.height="600%"}
multi_hist(metabo_measures[,MiMIR::metabolites_subsets$MET57])
```

## Binary Phenotypes
### Missingness of the binarized phenotypes
```{r missing binarized phenotypes, echo=FALSE}
  plot_na_heatmap(t(bin_phenotypes))
```

### Correlation of the binarized phenotypes
```{r corr pheno, echo=FALSE}
  res<-cor_assoc(data.matrix(bin_phenotypes),data.matrix(bin_phenotypes), bin_pheno_available, bin_pheno_available)
  plot_corply(res, main="Binarized phenotypes correlations", reorder.x=TRUE, abs=F, 
                      resort_on_p= TRUE,reorder_dend=F)
```

# Predicted variables
### Quality Control for the predicted values calculated
```{r quality control, echo=FALSE}
cat("The quality control settings chosen for MetaboAge are:\n")
a<-QCprep(as.matrix(metabo_measures[,MiMIR::metabolites_subsets$MET63]),
       MiMIR::PARAM_metaboAge,quiet=FALSE,
       Nmax_miss=Nmax_miss_metaboAge,
       Nmax_zero=Nmax_zero_metaboAge)

cat("The quality control settings chosen for the Surrogates are:\n")
b<- QCprep_surrogates(as.matrix(metabo_measures[,MiMIR::metabolites_subsets$MET63]),
                       MiMIR::PARAM_surrogates,quiet=FALSE,
                       Nmax_miss=Nmax_miss_surrogates,
                       Nmax_zero=Nmax_zero_surrogates)

```

### Missingness of the predicted values
```{r missing predicted values, echo=FALSE}
  plot_na_heatmap(t(predictors[,-1]))
```

### Histogram predictors
```{r histograms predicted scores, echo=FALSE, out.width="130%", out.height="400%"}
 multi_hist(predictors[,-1])
```

## Correlations of the predictors
```{r corr predictors, echo=FALSE}
res<-cor_assoc(predictors, predictors, colnames(predictors)[-1],colnames(predictors)[-1])
plot_corply(res, main="Correlations of the Metabolic scores", reorder.x=TRUE, abs=F,
                    resort_on_p= TRUE,reorder_dend=F)
```

# Accuracy
## ROC curves surrogates
```{r roc surrogates, echo=FALSE, out.width="150%", out.height="200%"}
    suppressWarnings(roc_surro_subplots(surrogates, bin_phenotypes))

```

## Surrogates t-test
```{r t-test surrogates, echo=FALSE}
   ttest_surrogates(surrogates = surrogates, bin_phenotypes = bin_phenotypes)
```

## Scatterplot MetaboAge
```{r metaboage scatterplot, echo=FALSE}
  x<-data.frame(phenotypes$age)
  rownames(x)<-rownames(phenotypes)
  scatterplot_predictions(x, MetaboAge,
                          title="Chronological Age vs MetaboAge",
                          xname="Chronological age",
                          yname="MetaboAge")
```

## Mortality Histogram divided in age ranges
```{r mort-age, echo=FALSE}
  hist_plots_mortality(mort_score,phenotypes)
```

# Calibration of the surrogates
## Reliability plots
```{r calibration, echo=FALSE, message=FALSE}
htmltools::tagList(lapply(bin_pheno_available, function(i){
    if(is.null(calibrations[[i]])){
    return(NULL)
  }else{
    surro<-calculate_surrogate_scores(met=metabo_measures, PARAM_surrogates = MiMIR::PARAM_surrogates,
                                    Nmax_miss=Nmax_miss_surrogates,
                                    Nmax_zero=Nmax_zero_surrogates,
                                    bin_names = MiMIR::phenotypes_names$bin_names,
                                    roc=F, quiet=T, post=F)
  surro<-surro$surrogates
    c1<-plattCalib_evaluation(r=bin_phenotypes[,i], p=surro[,paste0("s_",i)],
                          p.orig=surrogates[,paste0("s_",i)],name=paste0("s_",i), nbins = Nbins,
                          annot_x=c(1.22,1.22), annot_y=c(-0.9,-0.6))
    c1<-c1[c("cal.Plot","prob.hist")]
    subplot(c1, nrows = 2, shareX = F,shareY = F, titleX = F, titleY = F, which_layout = 1)
  }
 }))

```

## Correlations of the calibrated surrogates
```{r calibrations corr, echo=FALSE}
  calib<-calib_data_frame(calibrations, metabo_measures, bin_pheno_available)
  colnames(calib)<-paste0("s_",colnames(calib))
  res<-cor_assoc(calib,calib, colnames(calib),colnames(calib))
  plot_corply(res, main="Calibrated surrogates' correlations", reorder.x=TRUE, abs=F,
                      resort_on_p= TRUE,reorder_dend=F)
```


### Missingness of the calibrated surrogates
```{r missing calibrated, echo=FALSE}
  calib<-calib_data_frame(calibrations, metabo_measures, bin_pheno_available)
  plot_na_heatmap(t(calib))
```

### Sessino Info
```{r session info, echo=FALSE}
  sessionInfo()
```

